---
title: "Cleaning Gapminder Data"
author: "Tyler Byers"
date: "February 28, 2015"
output: html_document
---

This is my script for cleaning the Gapminder Data for the Data Vis project.  The data sets I'll combine are the GDP, Sugar Consumption, BMI, and Calories data sets

```{r load data}
library(ggplot2)
library(reshape2)
library(dplyr)
library(xlsx)
sugar <- read.xlsx('indicator sugar_consumption.xlsx',1)
bmi <- read.xlsx('Indicator_BMI male ASM.xlsx', 1)
gdp <- read.xlsx('GDPpercapitaconstant2000US.xlsx', 1)
cals <- read.xlsx('indicator food_consumption.xlsx', 1)
regions <- read.xlsx('Color- Regions.xlsx', 2)
```

###Clean Sugar Data

```{r, clean sugar}
sugar$X2004 <- NULL # 2004 data was a copy of 2003 data
colnames(sugar) <- substr(names(sugar), 2, 5)  # Remove the X from the Year
sugar$A..1 <- NULL # strange last column
sugar.t <- data.frame(t(sugar)) # Create a transposed data frame to be used later
names(sugar)[1] <- 'Country' # Rename column 1 into "country"
sugar.long <- melt(sugar, id.var = 'Country') # Melt into 3-variable dataframe
names(sugar.long)[2] <- 'Year' # Rename from "variable"
names(sugar.long)[3] <- 'Sugar Consumption' # Rename from "value"
```

###Clean BMI Data

```{r clean bmi}
colnames(bmi) <- substr(names(bmi), 2, 5)
names(bmi)[1] <- 'Country'
bmi.long <- melt(bmi, id.var = 'Country')
names(bmi.long)[2] <- 'Year'
names(bmi.long)[3] <- 'Body Mass Index'
```

###Clean GDP data

```{r clean gdp}
names(gdp) <- substr(names(gdp), 2, 5)
names(gdp)[1] <- 'Country'
gdp.long <- melt(gdp, id.var = 'Country')
names(gdp.long)[2] <- 'Year'
names(gdp.long)[3] <- 'GDP Per Capita'
```

### Clean Calorie data

```{r clean calories}
names(cals) <- substr(names(cals), 2, 5)
names(cals)[1] <- 'Country'
cals.long <- melt(cals, id.var = 'Country')
names(cals.long)[2] <- 'Year'
names(cals.long)[3] <- 'Calories Per Day'
```

### Clean Regions

```{r clean regions}
regions <- regions[,1:3]
names(regions) <- c('Country','Region','Digraph')
```

###Merge Data

```{r merge data}
sugar.bmi <- merge(sugar.long, bmi.long)
sugar.bmi.gdp <- merge(sugar.bmi, gdp.long)
cals.sugar.bmi.gdp <- merge(sugar.bmi.gdp, cals.long)
df_final <- merge(cals.sugar.bmi.gdp, regions)
cals.bmi <- merge(cals.long, bmi.long)
cals.bmi.gdp <- merge(cals.bmi, gdp.long)
cal_df_final <- merge(cals.bmi.gdp, regions)
```


###Add Square Root of GDP

When making charts in dimple, it appears as if the area of the bubbles for the GDP is not proportionate to GDP.  It looks like the RADIUS of the circles is, but that's not a good representation, since area of a circle is pi()*r^2.

```{r sqrt}
df_final$GDP.sqrt <- sqrt(df_final[,'GDP Per Capita'])
cal_df_final$GDP.sqrt <- sqrt(cal_df_final[,'GDP Per Capita'])
```

###Write Data

```{r write data}
# Write all data, with NAs
write.csv(df_final, file = 'cals_sugar_bmi_gdp.csv', row.names = FALSE)
# Write only data with complete cases
write.csv(df_final[complete.cases(df_final),], 
          file = 'compcases_cals_sugar_bmi_gdp.csv', row.names = FALSE)
# Write just cals data, complete cases only
write.csv(cal_df_final[complete.cases(cal_df_final),],
          file = 'calories_bmi_gdp.csv', row.names = FALSE)
```

